Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-07-18 articles/formalism/sections/introduction.tex
made minor revisions and finalized upto section 2.3

The pseudo-scalar $B$-modes particularly on large angular scales ($\sim 2$ degrees) are expected to have a significant contribution from primordial tensor perturbations generated during inflation. Hence a measurement of the large angular scale B-modes will yield information about the statistical properties of tensor perturbations which will finally lead to important constraints on models of inflation \cite{?}. The measurement of $B$-modes on small angular scales ($\sim$ few arcminute) will yield information on the clustering of matter across cosmic ages \cite{?}.

@@

Some of the ideas presented in this work bear resemblance to those in Zaldarriaga (2001) \citep{Zaldarriaga2001a}, however the approach adopted here is more rigorous and consequently there are updates to interpretation of the results and substantial differences in details.
\comment{``more rigorous'' is a little harsh}

This paper is organized in the following manner: In \sec{sec:pol-primer} we present a primer on the description of CMB polarization on the sphere, beginning with a heuristic argument that makes transparent the real space construction of $E$/$B$ modes.  We discuss the standard harmonic-space procedures for this operation. Finally, we introduce the matrix-vector notation which yields a more concise description of the harmonic space procedures. In \sec{sec:qu2eb} and \sec{sec:eb2qu} we derive and discuss the real space operators that transform $Q$/$U$ to $E$/$B$ and vice versa. In \sec{sec:visualize_operator} we evaluate these real space operators and present visualizations of these functions. In \sec{sec:purify_stokes_qu} we derive a real space operator that decomposes the Stokes $Q$/$U$  parameters into components that correspond to $E$ and $B$ modes respectively and present its visualizations.  In \sec{sec:radial_locality} we study the locality of the real space operators and explore its band limit dependence. In \sec{sec:generalized_operators} we present a systematic way to generalize the real space operators by controling the non-locality while recovering the standard power spectra, and we discuss the connection to the spin raising and lower operators. In \sec{sec:pol_filaments} we discuss $E/B$ mode signatures of foreground filaments. In \sec{sec:discussion}, we conclude with a summary and discuss the prospects this new method of analyzing CMB polarization.

>>>>

The pseudo-scalar $B$-modes particularly on large angular scales ($\sim 2$ degrees) are expected to have a significant contribution from primordial tensor perturbations generated during inflation. Hence a measurement of the large angular scale B-modes will yield information about the statistical properties of tensor perturbations which will finally lead to important constraints on models of inflation and the measurement of $B$-modes on small angular scales ($\sim$ few arcminute) will yield information on the clustering of matter across cosmic ages \citep{Abazajian2015, Kamionkowski2016,Abazajian2016,Hu2002c,Wehus2016}.

@@

\revisit{Some of the ideas presented in this work bear resemblance to those in Zaldarriaga (2001) \citep{Zaldarriaga2001a}, however the approach adopted here is more mathematically rigorous and consequently leads to updates in interpretation of the results and substantial differences in detail. This has enabled us to develop a real space polarization analysis tool, the details of which are beyond the scope of this article but will be discussed in a subsequent publication.}

This paper is organized in the following manner: In \sec{sec:pol-primer} we present a primer on the description of CMB polarization on the sphere, beginning with a heuristic argument that makes transparent the real space construction of $E$/$B$ modes.  We discuss the standard harmonic-space procedures for this operation. Finally, we introduce the matrix-vector notation which yields a more concise description of the harmonic space procedures. In \sec{sec:qu2eb} and \sec{sec:eb2qu} we derive and discuss the real space operators that transform $Q$/$U$ to $E$/$B$ and vice versa. In \sec{sec:visualize_operator} we evaluate these real space operators and present visualizations of these functions. In \sec{sec:purify_stokes_qu} we derive a real space operator that decomposes the Stokes $Q$/$U$  parameters into components that correspond to $E$ and $B$ modes respectively and present its visualizations.  In \sec{sec:radial_locality} we study the locality of the real space operators and explore its band limit dependence. In \sec{sec:generalized_operators} we present a systematic method of generalizing the real space operators by controlling the non-locality while recovering the standard power spectra and we discuss the connection to the standard spin raising and lower operators. In \sec{sec:pol_filaments} we discuss $E/B$ mode signatures of foreground filaments. In \sec{sec:discussion}, we conclude with a summary and discuss the prospects of this new method of analyzing CMB polarization maps.
